---
title: "RNAseq Pipeline Comparison"
author: "JD"
date: "2023-06-26"
output: html_document
---

## Introduction

This RMD contains scripts and figures to compare the van Heesch RNA-seq pipeline,
both for CentOS7 module style and for Rocky8 with apptainer containers

```{r libraries}
library(rtracklayer)
library(magrittr)
library(ggplot2)
```

```{r parameters}
wd="/hpc/pmc_vanheesch/projects/Jip/pilots/20230606_JD_containers/01_rnaseq"
```

## Stringtie

check stringtie GTF
```{r}
base_gtf <- rtracklayer::import(paste(wd,"analysis","base","stringtie",
                                      "BG96_HM1DCTRL_S13_L001","BG96_HM1DCTRL_S13_L001.gtf",sep="/")) 
container_gtf  <- rtracklayer::import(paste(wd,"analysis","container","stringtie",
                                      "BG96_HM1DCTRL_S13_L001","BG96_HM1DCTRL_S13_L001.gtf",sep="/")) 

length(base_gtf) - length(container_gtf)
```

For some reason, 35 rows in the GTF are different.

```{r }
length(unique(container_gtf$transcript_id))
length(unique(base_gtf$transcript_id))
```

There are 2 transcripts missing in the base GTF

## Stringtie merge

```{r}
base_merged_gffcompare_gtf <- rtracklayer::import(paste(wd,"analysis",
                                                        "base","gffcompare","base_merged","base_merged_gffcompare.annotated.gtf", sep = "/"))

container_merged_gffcompare_gtf <- rtracklayer::import(paste(wd,"analysis",
                                                        "container","gffcompare","container_pancreas","container_pancreas_gffcompare.annotated.gtf", sep = "/"))

length(base_merged_gffcompare_gtf) - length(container_merged_gffcompare_gtf)
```
Now we suddenly have a difference of 1400 lines in the Granges, where the base
has more far more lines. This is the biggest difference so far

```{r}
length(unique(base_merged_gffcompare_gtf$transcript_id))
length(unique(container_merged_gffcompare_gtf$transcript_id))
```

Now, we miss 76 transcripts that are not present in the container pipeline 
compared to the base pipeline

```{r}
base_novel_df <- as.data.frame(base_merged_gffcompare_gtf) %>%
  dplyr::filter(!(class_code == "="))
  
container_novel_df <- as.data.frame(container_merged_gffcompare_gtf) %>%
  dplyr::filter(!(class_code == "="))

table(base_novel_df$class_code)
table(container_novel_df$class_code)
```

## Filtered GTF

```{r}
base_filter_gtf <- rtracklayer::import(paste(wd,"analysis","base","customannotation",
                                             "base_merged_novel_filtered.gtf", sep = "/")
)

container_filter_gtf <- rtracklayer::import(paste(wd,"analysis","container","customannotation",
                                             "container_pancreas_novel_filtered.gtf", sep = "/")
)

length(base_filter_gtf) - length(container_filter_gtf)
```

Now, there is a different of 56 more lines in the container version (in 3 million, that is not a lot)

```{r}
length(unique(base_filter_gtf$transcript_id))

length(unique(container_filter_gtf$transcript_id))
```

the container version has 2 more TXs

```{r missing TXs}
base_diff <- setdiff(base_filter_gtf, container_filter_gtf)
container_diff <- setdiff(container_filter_gtf,base_filter_gtf)
```

In total, there are 5 ranges extra ranges in container DF and none in base diff. 
This seems small enough to disregard

```{r}
base_novel_df <- as.data.frame(base_filter_gtf) %>%
  dplyr::filter(!(class_code == "="))
base_novel_df <- as.data.frame(table(base_novel_df$class_code)) %>%
  dplyr::mutate(type = "base")

container_novel_df <- as.data.frame(container_filter_gtf) %>%
  dplyr::filter(!(class_code == "="))
container_novel_df <- as.data.frame(table(container_novel_df$class_code)) %>%
  dplyr::mutate(type = "container")

plot_df <- rbind(container_novel_df, base_novel_df)
colnames(plot_df) <- c("class_code","count","type")

base_df <- as.data.frame(table(base_novel_df$class_code)
table(container_novel_df$class_code)
```

```{r plot difference}

ggplot(plot_df, aes(x = class_code, y = count, fill = type,
                    group = type, label = count,
                    after_scale(alpha(fill, 0.4)))) +
  geom_bar(position = "dodge",stat = "identity") +
  geom_text(stat='identity', position = position_dodge(width = .9), vjust = -.3, size = 3.5) +
  scale_fill_manual(values = viridis::viridis(2, begin = .2, end = .6)) +
  theme_classic() +
  scale_y_log10() +
  theme(axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.y=element_blank())

ggsave(filename = paste(wd,"results","comparison_analysis","merged_annotated_gtf_comparison.pdf", sep ="/"))

```
